\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}

\setlength{\parindent}{0em}
\setlength{\parskip}{0.5em}

\title{CPM Vehicle Parameter Identification and Model Predictive Control}
\author{Janis Maczijewski}
\date{\today}

\begin{document}
\maketitle

\section{Vehicle Dynamics Model}


This is an end-to-end, grey-box model for the vehicle dynamics. The model parameters are not measured directly, but optimized to best fit the vehicle behavior.

\begin{align*}
\boldsymbol{x} &= [p_x, p_y, \psi, v] \\
\boldsymbol{u} &= [f, \delta, V] \\ 
\end{align*}



\begin{center}
\begin{tabular}{ r | l }
 $p_x$ & IPS x-position  \\ 
 $p_y$ & IPS y-position   \\ 
 $\psi$ & IPS yaw angle  \\ 
 $v$ & Odometer Speed  \\ 
 $f$ & Dimensionless motor command  \\ 
 $\delta$ & Dimensionless steering command  \\ 
 $V$ & Battery voltage  \\ 
\end{tabular}
\end{center}


\begin{align}
\dot{p}_x &= p_1 \cdot v \cdot (1+p_2 \cdot (\delta + p_{9})^2) \cdot \cos(\psi + p_3 \cdot (\delta + p_{9}) + p_{10}) \\
\dot{p}_y &= p_1 \cdot v \cdot (1+p_2 \cdot (\delta + p_{9})^2) \cdot \sin(\psi + p_3 \cdot (\delta + p_{9}) + p_{10}) \\
\dot{\psi} &= p_4 \cdot v \cdot (\delta + p_{9}) \\
\dot{v} &= p_5 \cdot v + (p_6 + p_7 \cdot V) \cdot \text{sign}(f) \cdot |f|^{p_8}
\end{align}

This is a kinematic bicycle model with some added terms to account for various errors.

\begin{itemize}
\item $p_1$: Compensate calibration error between IPS speed and odometer speed. 
\item $(1+p_2 \cdot (\delta + p_{9})^2)$: Compensate for speed differences due to different reference points between the IPS and odometer. The formulation is simplified with a second-order Taylor approximation.
\item $p_3$: Side slip angle (Schwimmwinkel) due to steering.
\item $p_{10}$: IPS Yaw calibration error.
\item $p_{4}$: Unit conversion for the steering state.
\item $p_{5}$: Speed low pass (PT1).
\item $p_{6}, p_{7}$: Motor strength depends on the battery voltage.
\item $p_{8}$: Compensate non-linear steady-state speed.
\item $p_{9}$: Steering misalignment correction.
\end{itemize}


\section{Model Discretization}

The model is discretized with the explicit Euler method, as follows:

\begin{align}
\boldsymbol{x}_{k+1} = \boldsymbol{x}_k + \Delta t \cdot f(\boldsymbol{x}_k,  \boldsymbol{u}_k, \boldsymbol{p}) 
\end{align}

This discretization is chosen for its simplicity and computational efficiency.

TODO justification inaccuracy: small timestep, and discretization is included in identification.

\section{Parameter Identification}

Optimal parameter estimation problem for the vehicle dynamics. The optimization tries to find a set of model parameters, that best explain/reproduce the experiment data.

\begin{align}
\underset{\boldsymbol{x}_k^j, \boldsymbol{p}}{\text{minimize}} && \sum_{j=1}^{n_{experiments}} \sum_{k=1}^{n_{timesteps}} E(\boldsymbol{x}_k^j - \hat{\boldsymbol{x}}_k^j) \\
\text{subject to} &&  \boldsymbol{x}_{k+1}^j = \boldsymbol{x}_k^j + \Delta t \cdot f(\boldsymbol{x}_k^j,  \hat{\boldsymbol{u}}_k^j, \boldsymbol{p}) \\
&& \quad k=1..(n_{timesteps}-1) \\
&& \quad j=1..n_{experiments} \\
\end{align} 



\begin{center}
\begin{tabular}{ r | l }
 $\hat{\boldsymbol{x}}_k^j$ & Measured States  \\ 
 $\hat{\boldsymbol{u}}_k^j$ & Measured Inputs   \\ 
 $f$ & Vehicle dynamics model  \\ 
 $\boldsymbol{p}$ & Model parameters  \\ 
 $\Delta t$ & Constant timestep $0.02s$ \\ 
 $E$ & Error penalty function \\ 
\end{tabular}
\end{center}

\textbf{Error penalty $E$}: Weighted quadratic error with model specific extensions. The yaw error function has a period of $2\pi$, so that a full rotation does not count as an error. This is done using $\sin(\Delta\psi/2)^2$.


\textbf{Delays}: This kind of optimization problem is not well suited for identifying the delay times (Totzeiten). The delays are solved in an outer loop. The delay is guessed/assumed and the measurement data is modified by appropriately shifting it in the time index $k$. This optimization problem is solved many times for combinations of delay times. The delays that create the lowest objective value are taken as the solution.


\section{Model Predictive Control}

The MPC uses the identified model to calculate control inputs in real time. The MPC must run in an environment with narrow computational constraints. It must run on a Raspberry Pi Zero W and in less than 10ms per time step. This results in a computational budget of roughly 100000 to 500000 double precision floating point operations per time step.


\subsection{Optimization Problem and Optimizer}

The MPC is formulated as a (mostly) unconstrained minimization problem, with a box constraint:

\begin{align}
\underset{z \in \mathbb{R}^n}{\text{minimize}} \quad & J(z) \\
\text{subject to} \quad & -1 \leq z_i  \leq 1 
\end{align}


The optimization problem is solved using a simple and lightweight method, the gradient descent method with momentum.
The method works by iterating the following two equations:

\begin{align*}
m^{(j)} &:= \beta  m^{(j-1)} - \nabla J(z^{(j-1)}) \\
z^{(j)} &:= \text{clip}(z^{(j-1)} + \alpha  m^{(j)})
\end{align*}

where
$z^{(j)}$ is the approximate solution, which is improved in each iteration,
$m^{(j)}$ is the momentum term,
$\alpha = 0.4$ and $\beta = 0.6$ are constants,
$\text{clip}(\cdot)$ applies the following function element-wise $\min(1, \max(-1, x_i))$ and
$\nabla J(z^{(j-1)})$ is the gradient of the objective.

A constant number of iterations $j = 1 ... N$ are executed. 
This guarantees a constant computation time, but not the convergence to the solution.
The approximate solution $z^{(N)}$ is however sufficient in practice.

\subsection{Trajectory Tracking Problem}

The goal of the MPC is to make the vehicle follow a given reference trajectory.
The reference trajectory is given as a Cubic Hermite spline function $r(t) = [p_{x,ref}, p_{y,ref}]$.
This function is evaluated on an appropriate time grid $t_k$ to give the discrete reference trajectory $[p_{x,ref,k}, p_{y,ref,k}]$ where $k = 1 ... H_p$.
The MPC objective is defined as follows:

\begin{align*}
J &= \sum_{k=1}^{H_p} \left[ (p_{x,k} - p_{x,ref,k})^2 + (p_{y,k} - p_{y,ref,k})^2 \right] \\
  &+ 0.5 \sum_{k=1}^{H_u}  (f_k - f_{k-1})^2 \\
  &+ 0.01 \sum_{m=1}^{H_u}  (\delta_{m} - \delta_{m-1})^2
\end{align*}

The predicted states are calculated explicitly and recursively. This is known as a single shooting method.

\begin{align*}
\begin{bmatrix}
p_{x,k+1}  \\
p_{y,k+1}  \\
\psi_{k+1}  \\
v_{k+1}  
\end{bmatrix}
= 
\begin{bmatrix}
p_{x,k}  \\
p_{y,k}  \\
\psi_{k}  \\
v_{k}  
\end{bmatrix}
+
\Delta t_{MPC} \cdot f\left(
\begin{bmatrix}
p_{x,k}  \\
p_{y,k}  \\
\psi_{k}  \\
v_{k}  
\end{bmatrix}
,  
\begin{bmatrix}
f_{\boldsymbol{m}_k}  \\
\delta_{\boldsymbol{m}_k}  \\
V_{\boldsymbol{m}_k} 
\end{bmatrix}
, \boldsymbol{p}\right), \quad k = 0 ... (H_p-1)
\end{align*}

where $f$ is the identified model from section (ref).

$\boldsymbol{m}$ is an index vector that allows reuse of input vectors. Current implementation $\boldsymbol{m} = [1, 1, 2, 2, 3, 3]$.

$H_p = 6$, $H_u = 3$

The MPC uses a longer discretization time step $\Delta t_{MPC} = 0.05s$. 

TODO experiment with the values for $\Delta t_{MPC}$, $H_p$, $H_u$.

The inputs for the trajectory tracking problem are:

\begin{itemize}
\item $[p_{x,0}, p_{y,0}, \psi_{0}, v_{0}]$: The measured state
\item $[f_{0}, \delta_{0}]$: The previous command
\item $[p_{x,ref,k}, p_{y,ref,k}]$: The discrete reference trajectory
\item $V_m$: The measured battery voltage
\item $\boldsymbol{p}$: The model parameters
\end{itemize}


The variables for the trajectory tracking problem are: $z = [f_1, ..., f_{H_u}, \delta_1, ..., \delta_{H_u}]$.


\subsection{Delay Compensation}


\subsection{CasADi and Code Generation}


\end{document}









